---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## loading libraries
```{r results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(lubridate)

```

## Loading and preprocessing the data
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")

summary(activity)
```

## What is mean total number of steps taken per day?
```{r}


df_activity_by_day <- activity %>% 
  group_by(date) %>%
  summarize(total_steps=sum(steps, na.rm = TRUE))

mean_total_steps_by_day <- mean(df_activity_by_day$total_steps)
median_total_steps_by_day <- median(df_activity_by_day$total_steps)


```

The mean of total number of steps taken by day is : `r  mean_total_steps_by_day` steps.
The median is : `r median_total_steps_by_day`.

```{r}
ggplot(df_activity_by_day, aes(total_steps)) +
  theme_minimal() +
  geom_histogram( binwidth = 500, stat = "bin", na.rm = TRUE)

```


## What is the average daily activity pattern?

```{r averagedailypattern}


df_activity_by_interval <- activity %>% 
  group_by(interval) %>%
  summarize(average_steps=mean(steps, na.rm = TRUE))
print(head(df_activity_by_interval))

max_average_steps_by_interval <- max(df_activity_by_interval$average_steps)
index_interval_max <- df_activity_by_interval$interval[which.max(df_activity_by_interval$average_steps)]

```

The max of average number of steps taken by interval is : `r  max_average_steps_by_interval` steps for the interval `r index_interval_max`


```{r}
ggplot(df_activity_by_interval, aes(interval)) +
  theme_minimal() +
  geom_line( aes(y= average_steps))

```


```{r}
ggplot(df_activity_by_interval, aes(average_steps)) +
  theme_minimal() +
  geom_histogram( binwidth = 5, stat = "bin", na.rm = TRUE)

```


## Imputing missing values

```{r}
na_count <- length(which(is.na(activity$steps)))
```

There is `r na_count` NAs value in the steps column.

Replace NA values with average step of the 5-min interval
```{r}

activity_removed_miss <- activity
v <- which(is.na(activity$steps))

find_interval_average <- function (x, df, column1, column2) { 
  index <- grep(pattern = paste("^",x,"$",sep=""), 
     df[[column1]], 
     value = FALSE
      )
  # print(head(list))
  return(df[index, column2])
  # return(column2)
}

interval_name <- activity$interval[v]
activity_removed_miss$steps[v] <- as.vector(sapply(interval_name, find_interval_average, df_activity_by_interval, "interval","average_steps"), mode = "numeric")

summary(activity_removed_miss)

```

Compute the total number of steps by day
```{r}
new_df_total_steps_by_day <- activity_removed_miss %>%
  group_by(date) %>%
  summarize(total_steps = sum(steps))

new_mean_total_steps_by_day  <- mean(new_df_total_steps_by_day$total_steps)
new_median_total_steps_by_day <- median(new_df_total_steps_by_day$total_steps)
```

The **mean** of total number of steps taken by day is : `r  new_mean_total_steps_by_day` steps (old value with NAs : `r  mean_total_steps_by_day`).
The **median** is : `r new_median_total_steps_by_day` (old value: `r median_total_steps_by_day`.)

**Histogram** of the total number of steps by day
```{r}
ggplot(new_df_total_steps_by_day, aes(x=total_steps))+
  theme_minimal() +
  geom_histogram(binwidth = 500)


```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity_removed_miss$weekday <- activity_removed_miss$date
activity_removed_miss$weekday <- (lubridate::wday(as.Date(activity_removed_miss$weekday),week_start = 1))

v <- activity_removed_miss$weekday == 6 | activity_removed_miss$weekday == 7
activity_removed_miss$weekday[v] <- "weekend"
activity_removed_miss$weekday[!v] <- "weekdays"
activity_removed_miss$weekday <- factor(activity_removed_miss$weekday)

summary(activity_removed_miss)

new_df_total_steps_by_interval <- activity_removed_miss %>%
  group_by(interval,weekday) %>%
  summarize(average_steps = mean(steps))

ggplot(new_df_total_steps_by_interval, aes(interval, colour = weekday )) +
  theme_minimal() +
  facet_wrap(. ~ weekday, nrow = 2) +
  theme(strip.text = element_text(size=18)) +
  geom_line( aes(y= average_steps))

```

Plot with time series combined

```{r}
ggplot(new_df_total_steps_by_interval, aes(interval, colour = weekday )) +

  theme_minimal() +
  geom_line( aes(y= average_steps))

```

